---
title: "COMPSCIX 415.2 Homework 3"
author: "Pankaj Mittal"
date: "June 21, 2018"
output:
  html_document: 
    self_contained: true
  pdf_document: default
---

My Github repository for my assignments can be found at this URL: https://github.com/pankajmit/compscix-415-2-assignments.git


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
library(ggplot2)
library(nycflights13)
```

## Section 3.8.1

### Q1) Problem with the plot - Some overlapping data has been hidden. 
This can be improved by adding position = "jitter"

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")
```

### Q2) width and height


### Q3) jitter adds small amount of random noise to show overlapping data. 
Count uses the size of the bubble to show hidden data. I liked the way count is 
showing the data and found it more useful

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter(width = 1)
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count()
```

### Q4) default is "dodge"

```{r}
base <- ggplot(data = mpg, mapping = aes(x=drv, y=hwy, colour=class)) 
base + geom_boxplot()
base + geom_boxplot(position='dodge')  #This is same as above
base + geom_boxplot(position='identity')  #This is different and overlapping
base + geom_boxplot(position='jitter')   #This is different and confusing
```


## Section 3.9.1

### Q2) Labs is used to modify axis, legend and labels

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot() +
  coord_flip() +
  labs(y = "Highway MPG", x = "", title = "Highway MPG by car class")
```

### Q4) cty and hwy are related directly propsotional.
coord_fixed is important to ensure that x and y axix are using the same units for proper comparison
geom_abline adds the reference line to give visual representation of trends in data

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline(intercept = 0) +
  coord_fixed()
```


## Section 4.4
### Q1) variable name (my_var1able) has been misspelled

### Q2) Fixed code
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
filter(mpg, cyl == 8)
filter(diamonds, carat > 3)
```

## Section 5.2.4
### Q1) Run flight queries
```{r}
filter(flights, arr_delay > 120) # arr_delay greater than 2 hours
filter(flights, dest %in% c("IAH", "HOU")) # Flights from Houston
filter(flights, carrier %in% c("UA", "AA", "DL")) # United, American and Delta
filter(flights, month %in% c(7,8,9)) # Month of July to Sep
filter(flights, arr_delay > 120, dep_delay < 1) # Arrived late but departed on time
filter(flights, dep_delay >= 60 , (dep_delay - arr_delay) > 30) # covered atleast 30 minutes in air
filter(flights, between(dep_time, 0, 600)) # Early morning departure
```

### Q3) 8,255 flights
```{r}
filter(flights, is.na(dep_time))
```
Other missing variables are related to actual departure and arrival. 
This data may be related to future travel dates

### Q4) Generic rule for NA handling
```{r}
NA ^ 0 #Anything with zero exponent is always 1 (irrespective of base)
NA | TRUE #Anything when OR with TRUE will always be TRUE (irrespectine of other variable)
FALSE & NA #Anything when AND with FALSE will always be FALSE (irrespectine of other variable)
```

## Section 5.4.1
### Q1) Selection of columns in various ways

```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, c(dep_time, dep_delay, arr_time, arr_delay))
select(flights, 4,6,7,9)
select(flights, starts_with("dep_"), starts_with("arr_"))
```

### Q3) one_of() is used to list all the needed variables in character vector.

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, vars)
```

This will be useful to select various columns from the dataset

