---
title: "COMPSCIX 415.2 Homework 6"
author: "Pankaj Mittal"
date: "July 11, 2018"
output:
  html_document: 
    self_contained: true
  pdf_document: default
---

My Github repository for my assignments can be found at this URL: https://github.com/pankajmit/compscix-415-2-assignments.git


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
library(ggplot2)
library(nycflights13)
library(forcats)
library(lubridate)
library(mosaicData)
```

## Exercise 1

### 1. Variables in this data set are - outcome (alive/dead), smoker (yes/no) and age 
### 2. There are 1,314 observation. "outcome" represents whether the person is alive or dead after 20 years;  "Smoker" represents smoking status at the baseline; "age" represents age at the time of first survey
### 3. Data is shown below but it does not convey an meaningful information, because age is missing

```{r}
Whickham %>% count(outcome, smoker)
ggplot(data = Whickham, mapping = aes(x = outcome, y = smoker)) + 
  geom_point()
```

### 4. Results are not expected for the age group > 64. More people are dead who are non-smoker. This could be result of natural death of old age

```{r}

Whickham_age_factor <- Whickham %>%
  mutate(age_factor = case_when(
                        age <= 44 ~ "age <= 44", 
                        age > 44 & age <= 64 ~ "age > 44 & age <=64", 
                        age > 64 ~"age > 64"))
Whickham_age_count <- Whickham_age_factor %>%
  group_by(outcome, smoker, age_factor) %>%
  summarize(age_count = n())

ggplot(data = Whickham_age_count, mapping = aes(x = outcome, y = smoker, size = age_count)) + 
  geom_point() + 
  facet_wrap(~age_factor)
```

## Exercise 2

### 1. Histogram and density curve of gamma distribution 

```{r}
n <- 10000
gamma_samp <- tibble(x = rgamma(n, shape = 1, scale = 2))
ggplot(data = gamma_samp, mapping = aes(x)) + 
  geom_histogram()
ggplot(data = gamma_samp, mapping = aes(x)) + 
  geom_density()
```

### 2. mean = 1.983, standard deviation = 1.957

```{r}
mean(gamma_samp$x)
sd(gamma_samp$x)
```

### 3. mean = 2.082, standard deviation = 2.014

```{r}
n <- 30
gamma_samp <- tibble(x = rgamma(n, shape = 1, scale = 2))
ggplot(data = gamma_samp, mapping = aes(x)) + 
  geom_histogram()
ggplot(data = gamma_samp, mapping = aes(x)) + 
  geom_density()
mean(gamma_samp$x)
sd(gamma_samp$x)
```

### 4. Build the vector of mean samples with 10000 entries

```{r}
mean_samp <- rep(NA, 10000)
for(i in 1:10000) {
  g_samp <- rgamma(30, shape = 1, scale = 2)
  mean_samp[i] <- mean(g_samp)
}

mean_samp <- tibble(mean_samp)

```

### 5. Build histogram of collection of means

```{r}
ggplot(data = mean_samp, mapping = aes(mean_samp)) + 
  geom_histogram()
```

### 6. mean = 2.005, SD = 0.366

```{r}
mean(mean_samp$mean_samp)
sd(mean_samp$mean_samp)
```

### 7. Any surprize - Standard deviation is very small compared to gamma distribution


### 8. Results are very close but this could be even closer if sample size was bit bigger

```{r}

mean_samp <- rep(NA, 300)
for(i in 1:300) {
  g_samp <- rgamma(30, shape = 1, scale = 2)
  mean_samp[i] <- mean(g_samp)
}

mean_samp <- tibble(mean_samp)

ggplot(data = mean_samp, mapping = aes(mean_samp)) + 
  geom_histogram()

mean(mean_samp$mean_samp)
sd(mean_samp$mean_samp)
```


