
---
title: "COMPSCIX 415.2 Homework 7"
author: "Pankaj Mittal"
date: "July 20, 2018"
output:
  html_document: 
    self_contained: true
  pdf_document: default
---

My Github repository for my assignments can be found at this URL: https://github.com/pankajmit/compscix-415-2-assignments.git


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
library(ggplot2)
library(broom)
```


## Exercise 1
### Number of observations:1,460 ; Number of columns: 81
```{r}
file_path_read <- "C:/training/train.csv"
houses <- read_csv(file = file_path_read)
glimpse(houses)
```

## Exercise 2 
### Visulaize the distribution of SalePrice
```{r}
ggplot(data = houses) +
  geom_histogram(mapping = aes(x = SalePrice), binwidth = 10000)
```

### Visualize the covariation between SalePrice and (Neighborhood, OverallQual)
```{r}
ggplot(houses) +
  geom_point(mapping = aes(x=SalePrice, y = Neighborhood))

ggplot(houses) +
  geom_point(mapping = aes(x=SalePrice, y = OverallQual))

```

## Exercise 3
### Coefficient = 180921; Coeffiecient is same as average value of SalePrice; R-Squared value is zero. This means this model is not explaning any variability.

```{r}
houses_lm <- lm(formula = SalePrice ~ 1, data = houses)
tidy(houses_lm)
summary(houses_lm)
glance(houses_lm)
```

## Exercise 4
```{r}
houses_lm <- lm(formula = SalePrice ~ GrLivArea + OverallQual + Neighborhood, data = houses)
summary(houses_lm)
```

### How would you interpret the coeﬃcients on GrLivArea and OverallQual?
#### Coefficient of GrLivArea = 55.56; i.e. For every 1 unit increase in living area, Sales Price increase by about $55.
#### Coefficient of OverallQual = 20951.42; i.e. For every 1 unit increase in overall quality, Sales Price increase by about $10,951. Please note this attribute has only 10 possible values.

### How would you interpret the coeﬃcient on NeighborhoodBrkSide?
#### coeﬃcient of NeighborhoodBrkSide is -13025.45. This means, average price difference between "Brookside"" and "Bloomington Heights" neighborhood is -($13,035)

### Are the features signiﬁcant?
#### Looking at the p-value, GrLivArea and OverallQual seems significant.
#### Though, NeighborhoodBrkSide is not significant as p-value is about 0.21.

### Are the features practically signiﬁcant?
#### It is common knowedge that houses with greater living area and better quality would have higher price. These variables are significant bit have limited value in adding the knowledge through data science

### Is the model a good ﬁt?
#### With adjusted R square value of 0.783, model is a good fit

## Exercise 6
### Coefficient of X and the R-squared values are changing with lot of variability

```{r}
sim1a <- tibble( x = rep(1:10, each = 3), 
                    y = x * 1.5 + 6 + rt(length(x), df = 2) )
sim1a
sim1a_lm <- lm(formula = y ~ x, data = sim1a)
summary(sim1a_lm)
```



