<<<<<<< HEAD

---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Pankaj Mittal"
date: "July 5, 2018"
output:
  html_document: 
    self_contained: true
  pdf_document: default
---

My Github repository for my assignments can be found at this URL: https://github.com/pankajmit/compscix-415-2-assignments.git


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(mdsr)
library(ggplot2)
library(tidyverse)
library(nycflights13)
```

## Table of Contents
1. RStudio and R Markdown
2. The tidyverse packages
3. R Basics
4. Data Import / Export
6. Visualization
6. EDA

## R Studio and R Markdown
### Each midterm question is clearly separated in this document with table cof content shown above

## The tidyverse packages

### Question 1: Name which package is associated with each task below?
#### Plotting - ggplot2
#### Data munging/wrangling - dplyr
#### Reshaping Data - tidyr
#### Importing/Exporting data - readr

### Question 2: Name two functions from each package?
#### Plotting (ggplot2) - geom_point, geom_boxplot
#### Data munging/wrangling (dplyr) - select, filter, summarise, mutate
#### Reshaping Data(tidyr) - gather, spread 
#### Importing/Exporting data (readr) - read_csv, read_delim, write_csv

## R Basics

### Question 1: Fix the code with fewest number of possible changes?
#### Corrected code is below:
```{r}
My_data.name___is.too00ooLong <- c( 1 , 2 , 3 )
My_data.name___is.too00ooLong
```

### Question 2: Fix the code so it works?
#### Corrected code is below
```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
my_string
```

### Question 3: Comment on what happened to the vector?
#### Answer: All entries for the vector were forced to convert to character
```{r}
my_vector <- c(1, 2, '3', '4', 5)
my_vector
is.character(my_vector)
is.double(my_vector)
```

## Data import/export

### Question 1: Showcase import of rail_trail.txt file

```{r}
file_path_read <- "C:/training/rail_trail.txt"
x <- read_delim(file = file_path_read, delim = "|")
glimpse(x)
```

### Question 2: Showcase export of rail_trail.txt file

```{r}
file_path_write <- "C:/training/rail_trail.csv"
write_csv(x, path = file_path_write)
y <- read_csv(file = file_path_write)
glimpse(y)
```


## Visualization

### Question 1: Critique this graphic, giving three examples
#### 1. This data is better represented as bar chart rather than bubble chart
#### 2. Percentages are not adding up to 100. Those should have been explained
#### 3. It would have been better to see the age breakup for men and women rather than completely separate visual 

### Question 2: Reproduce the graphic shown using *diamonds* data set

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, y = carat)) + 
  geom_boxplot() +
  labs(x = "CUT OF DIAMOND", y = "CARAT OF DIAMOND") +
  coord_flip()
```


### Question 3: How to make above graphic more useful
#### Answer: This graphic could be more useful if box plot is drawn separately for each color, rarther than putting together. 

```{r}
ggplot(data = diamonds, mapping = aes(x = reorder(cut, carat, FUN = median), y = carat, fill=color)) + 
  geom_boxplot() +
  labs(x = "CUT OF DIAMOND", y = "CARAT OF DIAMOND") +
  coord_flip() +
  facet_wrap(~color)
```


## Data Munging and Wrangling

### Question 1: Is this data tidy?
#### No, this data is not tidy. Below code will make this tidy.
```{r}
table2 %>%
  spread(key = type, value = count)
```

### Question 2: Create a new column called price_per_carat? Do not print the results
#### Here is the code
```{r echo = TRUE, results = FALSE}
diamonds %>%
  mutate(price_per_carat = price / carat)
```

### Question 3: How many diamond (and what proportion) have price > 10000 and a carat < 1.5?
#### Answer: 834 diamonds with about 1.54% proportion

```{r}
diamonds %>%
  mutate(selected = (price > 10000 & carat < 1.5)) %>%
  group_by(selected) %>%
  summarise(n = n()) %>%
  mutate(prop = n/sum(n))
```

#### These results does makes sense. Number is small because smaller diamonds (carat < 1.5) are geenrally costlier than 10000
#### We just need to be careful of NULL values in the data

## Exploratory Data Analysis (txhousing data)

### Question 1: During what time period is the data from?
#### Answer: From: Jan 2000 To: Dec 2015
```{r}
txhousing %>%
  summarize(min_year = min(year), min_mo = min(month), max_year = max(year), max_mo = max(month))
```

### Question 2: How many cities are represented?
#### Answer: 46 cities are represented

```{r}
txhousing %>%
  count(city)
```

### Question 3: Which city, month and year had the highest number of sales?
#### Answer: City = Houston, Month = 7 (July), Year = 2015
```{r}
txhousing %>%
  filter(sales == max(sales, na.rm = TRUE))
```

### Question 4: What kind of relationship exists between the number of listings and the number of sales?
#### Answer: In general, as number of listings increases, number of sales is also increasing

```{r}
ggplot(data = txhousing, aes(sales, listings)) +
  geom_point()
```


### Question 5: What proportion of sales is missing for each city?
#### Answer: Please see the results of query below

```{r}
txhousing %>%
  mutate(missed_sales = is.na(sales)) %>%
  group_by(city, missed_sales) %>%
  summarise(count = n()) %>%
  mutate(prop = count / sum(count))
```


### Question 6: Look at only the cities and months with greater than 500 sales
```{r}
txhousing_over500 <- txhousing %>%
  filter(sales > 500) 
txhousing_over500
```

#### 6)a) Are the distributions of the median sales price, when grouped by city, different?
#### Answer: distribution of median sales price is different as per chart below
```{r}
ggplot(data = txhousing_over500, mapping = aes(x = city, y = median)) + 
  geom_point() 
```


#### 6)b) Any cities that stand out that you’d want to investigate further?
#### Answer: Houston has wide range of sales; Dallas has too many ouliers on the higher median price
```{r}
ggplot(data = txhousing_over500, mapping = aes(x = city, y = median)) + 
  geom_boxplot() +
  coord_flip() 

ggplot(data = txhousing_over500, mapping = aes(x = city, y = sales)) + 
  geom_point() +
  coord_flip() 

```


#### 6)c) Why might we want to filter out all cities and months with sales less than 500?
#### Answer: This makes the data set targeted for our analysis without the need of any filtering
=======

---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Pankaj Mittal"
date: "July 5, 2018"
output:
  html_document: 
    self_contained: true
  pdf_document: default
---

My Github repository for my assignments can be found at this URL: https://github.com/pankajmit/compscix-415-2-assignments.git


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(mdsr)
library(ggplot2)
library(tidyverse)
library(nycflights13)
```

## Table of Contents
1. RStudio and R Markdown
2. The tidyverse packages
3. R Basics
4. Data Import / Export
6. Visualization
6. EDA

## R Studio and R Markdown
### Each midterm question is clearly separated in this document with table cof content shown above

## The tidyverse packages

### Question 1: Name which package is associated with each task below?
#### Plotting - ggplot2
#### Data munging/wrangling - dplyr
#### Reshaping Data - tidyr
#### Importing/Exporting data - readr

### Question 2: Name two functions from each package?
#### Plotting (ggplot2) - geom_point, geom_boxplot
#### Data munging/wrangling (dplyr) - select, filter, summarise, mutate
#### Reshaping Data(tidyr) - gather, spread 
#### Importing/Exporting data (readr) - read_csv, read_delim, write_csv

## R Basics

### Question 1: Fix the code with fewest number of possible changes?
#### Corrected code is below:
```{r}
My_data.name___is.too00ooLong <- c( 1 , 2 , 3 )
My_data.name___is.too00ooLong
```

### Question 2: Fix the code so it works?
#### Corrected code is below
```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
my_string
```

### Question 3: Comment on what happened to the vector?
#### Answer: All entries for the vector were forced to convert to character
```{r}
my_vector <- c(1, 2, '3', '4', 5)
my_vector
is.character(my_vector)
is.double(my_vector)
```

## Data import/export

### Question 1: Showcase import of rail_trail.txt file

```{r}
file_path_read <- "C:/training/rail_trail.txt"
x <- read_delim(file = file_path_read, delim = "|")
glimpse(x)
```

### Question 2: Showcase export of rail_trail.txt file

```{r}
file_path_write <- "C:/training/rail_trail.csv"
write_csv(x, path = file_path_write)
y <- read_csv(file = file_path_write)
glimpse(y)
```


## Visualization

### Question 1: Critique this graphic, giving three examples
#### 1. This data is better represented as bar chart rather than bubble chart
#### 2. Percentages are not adding up to 100. Those should have been explained
#### 3. It would have been better to see the age breakup for men and women rather than completely separate visual 

### Question 2: Reproduce the graphic shown using *diamonds* data set

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, y = carat)) + 
  geom_boxplot() +
  labs(x = "CUT OF DIAMOND", y = "CARAT OF DIAMOND") +
  coord_flip()
```


### Question 3: How to make above graphic more useful
#### Answer: This graphic could be more useful if box plot is drawn separately for each color, rarther than putting together. 

```{r}
ggplot(data = diamonds, mapping = aes(x = reorder(cut, carat, FUN = median), y = carat, fill=color)) + 
  geom_boxplot() +
  labs(x = "CUT OF DIAMOND", y = "CARAT OF DIAMOND") +
  coord_flip() +
  facet_wrap(~color)
```


## Data Munging and Wrangling

### Question 1: Is this data tidy?
#### No, this data is not tidy. Below code will make this tidy.
```{r}
table2 %>%
  spread(key = type, value = count)
```

### Question 2: Create a new column called price_per_carat? Do not print the results
#### Here is the code
```{r echo = TRUE, results = FALSE}
diamonds %>%
  mutate(price_per_carat = price / carat)
```

### Question 3: How many diamond (and what proportion) have price > 10000 and a carat < 1.5?
#### Answer: 834 diamonds with about 1.54% proportion

```{r}
diamonds %>%
  mutate(selected = (price > 10000 & carat < 1.5)) %>%
  group_by(selected) %>%
  summarise(n = n()) %>%
  mutate(prop = n/sum(n))
```

#### These results does makes sense. Number is small because smaller diamonds (carat < 1.5) are geenrally costlier than 10000
#### We just need to be careful of NULL values in the data

## Exploratory Data Analysis (txhousing data)

### Question 1: During what time period is the data from?
#### Answer: From: Jan 2000 To: Dec 2015
```{r}
txhousing %>%
  summarize(min_year = min(year), min_mo = min(month), max_year = max(year), max_mo = max(month))
```

### Question 2: How many cities are represented?
#### Answer: 46 cities are represented

```{r}
txhousing %>%
  count(city)
```

### Question 3: Which city, month and year had the highest number of sales?
#### Answer: City = Houston, Month = 7 (July), Year = 2015
```{r}
txhousing %>%
  filter(sales == max(sales, na.rm = TRUE))
```

### Question 4: What kind of relationship exists between the number of listings and the number of sales?
#### Answer: In general, as number of listings increases, number of sales is also increasing

```{r}
ggplot(data = txhousing, aes(sales, listings)) +
  geom_point()
```


### Question 5: What proportion of sales is missing for each city?
#### Answer: Please see the results of query below

```{r}
txhousing %>%
  mutate(missed_sales = is.na(sales)) %>%
  group_by(city, missed_sales) %>%
  summarise(count = n()) %>%
  mutate(prop = count / sum(count))
```


### Question 6: Look at only the cities and months with greater than 500 sales
```{r}
txhousing_over500 <- txhousing %>%
  filter(sales > 500) 
txhousing_over500
```

#### 6)a) Are the distributions of the median sales price, when grouped by city, different?
#### Answer: distribution of median sales price is different as per chart below
```{r}
ggplot(data = txhousing_over500, mapping = aes(x = city, y = median)) + 
  geom_point() 
```


#### 6)b) Any cities that stand out that you’d want to investigate further?
#### Answer: Houston has wide range of sales; Dallas has too many ouliers on the higher median price
```{r}
ggplot(data = txhousing_over500, mapping = aes(x = city, y = median)) + 
  geom_boxplot() +
  coord_flip() 

ggplot(data = txhousing_over500, mapping = aes(x = city, y = sales)) + 
  geom_point() +
  coord_flip() 

```


#### 6)c) Why might we want to filter out all cities and months with sales less than 500?
#### Answer: This makes the data set targeted for our analysis without the need of any filtering
>>>>>>> 4ce815d02d5617e9a19a808963b9ad3315f65471
